```{r setup, include=FALSE} knitr::opts_chunk$set(echo = TRUE)
library(knitr) 
library(ggplot2)
source("analysis.R")
```
## Introduction
Racial disparities in U.S. incarceration rate has been a topic of discussion. We have been told that communities of color make up a disproportionate 
share of the American imprisoned population, especially Black people. By analyzing the incarceration trends data collected by the Vera Institute of Justice,
this report investigates the jail and prison population counts of black and white people across the U.S. to learn about the black-to-white incarceration ratio 
and incarceration counts of black individuals. <br>
According to the data, **`r state_highest_prison_race_ratio`** (Louisiana) has the highest black-to-white incarceration ratio for both jail (2.26) and prison 
(3.17) population, (excluding `r state_highest_jail_race_ratio`). In addition, 11 states have a ratio greater than 1.0 for the jail population, and 17 states have
a ratio greater than 1.0 for the prison population. In terms of the total incarcerated black population, **`r state_highest_black_incarceration`** (California) is in
the first place with a number of **`r ca_black_incarceration`**. However, incarceration counts of black people have declined. With the most recent data available,
the jail incarceration count declined by **`r black_jail_change_year`** between 2008 and 2018, and the prison incarceration count
declined by **`r black_prison_change_year`** between 2006 and 2016. <br>

## Chart 1
```{r, echo=FALSE}
trends_over_time
```
**(Note that only jail incarceration counts were collected in 2017 and 2018.)** <br>
This chart is included to look into an interesting pattern in the total incarcerated (jail and prison) black population in different states. While California 
has the highest number of total accumulated incarcerated black people, it is not the case when examining year to year, according to the chart. 
The number in **Texas** surged in 2015 and has surpassed California in most of the following years. Why there was a surge and whether Texas would become the
state where the number of incarcerated black population is the highest are areas worth digging into. <br>
## Chart 2
```{r, echo=FALSE}
variable_comparison
```
This line graph is included to compare the incarceration count (jail and prison) between black and white people throughout the years. 
Before analyzing the data, I expected the number of incarcerated black people to be much higher than that of the white since racial discrimination is 
believed to have been ingrained in the criminal legal and law enforcement system. However, according to the graph, the number of incarcerated white people
surpassed that of the black around 2010. This result also corresponded to the finding that incarceration counts of black people have declined.
Racial injustice has not gone away, but I wonder sometimes we may be too biased before exploring the data.
## Map
```{r, echo=FALSE}
map
```
<br>
This graph depicts the severity of the black-to-white incarceration ratio in the United States over the previous several decades. Between 2006 and 2007, both have an identical number of incarcerations in the United States, but after 2010, the number of white incarcerations begins to exceed the number of black incarcerations.
